---
title: "DATA607-Assignment1"
author: "N Nedd"
date: "February 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("RCurl")
library(RCurl)
#install.packages("plyr")
library(plyr)
```

#Load Dataset

```{r load-dataset, eval = TRUE}
location <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"

mushroomData <- getURL(location)
mushroomDF <-  read.csv(text=mushroomData,header=F,sep=",")
head(mushroomDF)
```

#Extract relevant columns

**The columns corresponding to cap-color, odor, gill-color, spore-print-color will be selected.  Since the first column is the indicator of poisonous/edible this will translate to columns 1,4,6,10,21

```{r extract-cols, eval = TRUE}
mushroomDF <- mushroomDF[,c(1,4,6,10,21)]
str(mushroomDF)
```

#Change column names:

```{r change-colnames, eval = TRUE}
colnames(mushroomDF) <- c("edibility", "cap-color", "odor", "gill-color","spore-print-color")
head(mushroomDF)
```
#change letters to descriptive names

**According to the data description found at the following link: (https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names) the abbreviations are as follows:**

1. cap-color: brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
2. odor: almond=a,anise=l,creosote=c,fishy=y,foul=f,musty=m,none=n,pungent=p,spicy=s
3. gill-color: black=k,brown=n,buff=b,chocolate=h,gray=g,green=r,orange=o,pink=p,purple=u,red=e,white=w,yellow=y
4. spore-print-color: black=k,brown=n,buff=b,chocolate=h,green=r,orange=o,purple=u,white=w,yellow=y


```{r descriptive-values, eval = TRUE}
mushroomDF$edibility <- revalue(mushroomDF$edibility, c("e" = "edible", "p" = "poisonous"))
mushroomDF$`cap-color` <- revalue(mushroomDF$`cap-color`,c("n" = "brown", "b" = "buff", "c" = "cinnamon",
                                                           "g" = "gray", "r" = "green", "p" = "pink",
                                                           "u" = "purple", "e" = "red", "w" = "white",
                                                           "y" = "yellow"))
mushroomDF$odor<- revalue(mushroomDF$odor, c("a" = "almond", "l" = "anise", "c" = "creosote",
                                                           "y" = "fishy", "f" = "foul", "m" = "musty",
                                                           "n" = "none", "p" = "pungent", "s" = "spicy"))

mushroomDF$`gill-color`<- revalue(mushroomDF$`gill-color`, c("k" = "black", "n" = "brown", "b" = "buff", 
                                                             "h" = "chocolate", "g" = "gray", "r" = "green",                                                              "o" = "orange", "p" = "pink", "u" = "purple", 
                                                             "e" = "red", "w" = "white",
                                                             "y" = "yellow"))
mushroomDF$`spore-print-color` <- revalue(mushroomDF$`spore-print-color`, 
                                                            c("k" = "black", "n" = "brown", "b" = "buff", 
                                                             "h" = "chocolate", "r" = "green",                                                                            "o" = "orange", "u" = "purple", "w" = "white",                                                               "y" = "yellow"))
```

#Display details of transformed data frame

```{r display-dataset,eval = TRUE}
str(mushroomDF)

head(mushroomDF)
```


