---
title: "Data 607 - Project 1"
author: "N Obieyisi and N Nedd"
date: "February 26, 2017"
output: html_document
---

#Setup packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if("stringr" %in% rownames(installed.packages()) == FALSE) {install.packages("stringr")}
library(stringr)

```

#Read in Data

```{r readin, eval = TRUE, warning = FALSE}
raw_data <- scan("https://raw.githubusercontent.com/NNedd/CUNYMSDA/master/Data607/Project1/tournamentinfo.txt", what = "character", skip = 4,sep = "|")
head(raw_data, 23)

tourn_all <- "https://raw.githubusercontent.com/NNedd/CUNYMSDA/master/Data607/Project1/tournamentinfo.txt"

tlines <- readLines(tourn_all)
```

#Format data into data frame
For examination of the data the first record starts in row 1 and ends in row 23 (the 23rd row being ------)
Therefore the records will start in the following rows: 1, 24, 47, 70, etc. 


```{r makeDF, eval = TRUE}
format_data <- NULL  #setup data frame
i <- 1
while( i < length(raw_data)) {
  Number <- as.numeric(raw_data[i])  
  Name <- raw_data[i+1]
  Total<- as.numeric(raw_data[i+2])
  Round1 <- as.numeric(str_extract(raw_data[i+3], "[[:digit:]]+"))
  Round2 <- as.numeric(str_extract(raw_data[i+4], "[[:digit:]]+"))
  Round3 <- as.numeric(str_extract(raw_data[i+5], "[[:digit:]]+"))
  Round4 <- as.numeric(str_extract(raw_data[i+6], "[[:digit:]]+"))
  Round5 <- as.numeric(str_extract(raw_data[i+7], "[[:digit:]]+"))
  Round6 <- as.numeric(str_extract(raw_data[i+8], "[[:digit:]]+"))
  Round7 <- as.numeric(str_extract(raw_data[i+9], "[[:digit:]]+"))
  Player_state <- raw_data[i+11]
  
  ranking_info <- str_replace_all(raw_data[i+12], " ","")
  player_rank <- as.numeric(str_replace(str_extract(ranking_info, "[:][[:digit:]]{3,4}"), ":", ""))
  
  format_data<- rbind(format_data, data.frame(Number, Name, Total, Round1, Round2, Round3, Round4, Round5, Round6, Round7, Player_state, player_rank))
  i <- i +23
  
}

head(format_data)
```


#Calculate Average Pre-tournament Rating of Opponents
```{r averages, eval = TRUE}
Rowlen <-nrow(format_data)
i <- 1
for (i in 1:Rowlen )
{
  opponents <- as.numeric(format_data[i,c(4:10)]) #store the opponent playerIds 
  
  Avg_counter <- 0
  Sum <- 0
  for (j in 1:7)
  {
    
    #Ensure to perform average only for the correct number of games played
    if(is.na(format_data[opponents[j],12]) == FALSE ) 
    {
      Sum <- Sum + format_data[opponents[j],12] 
      
      Avg_counter <- Avg_counter + 1
    }
    
  }
  
  Avg_Pre_chess_rating <- Sum/Avg_counter
  
  format_data[i,13] <-Avg_Pre_chess_rating
  

  
}
```

#Clean Up dataframe for output

```{r clean-up, eval = TRUE}
#Select only needed columns
col_index <- c(2,11,3,12,13)
tournament_data <- format_data[,col_index]

#Cleanup column names
colnames(tournament_data) <- c("Player's Name", "Player's State", "Total Number of Points", "Player's Pre Rating", "Average Pre Chess Rating of Opponents")

#Cleanup data
tournament_data[,1] <- str_trim(tournament_data[,1], side = "both")
tournament_data[,2] <- str_trim(tournament_data[,2], side = "both")
tournament_data[,5] <- round(tournament_data[,5])


```

#Create CSV

```{r output-csv, eval = TRUE}
#write to csv
write.table(tournament_data, file = "tournamentData.csv",row.names=FALSE, na="",col.names=TRUE, sep=",")

```

#Calcuate for each player, number of wins, draws and loss

##Collect Win, loss, draw data
```{r}



all <- c(seq(1,length(tlines), by=1)) #all row numbers
skip <- c(seq(1,length(tlines), by=3)) #row numbers to skip
sub_rows <- !(all %in% skip) #filter out rows to skip from all rows
tourn_all_chess <- tlines[c(sub_rows)]

tourn_all_chess_list <-str_split(string=tourn_all_chess,pattern="[|]", n = Inf, simplify = FALSE)

n<-length(tourn_all_chess)
tourn_all_chess_df <- data.frame(matrix(unlist(tourn_all_chess_list), nrow=n, byrow=T),stringsAsFactors=FALSE)

n <- nrow(tourn_all_chess_df ) #rows of unstructured df
r1 <- seq(2,n,2) #row sequence of attributes to be pivoted to columns
r2 <- seq(1,n-1,2) #row sequence of attributes to be kept as rows
tourn_all_sub <- tourn_all_chess_df[r2,] #subset of data frame per player name


for (i in 1:10)
{
 tourn_all_sub<- cbind(tourn_all_sub,tourn_all_chess_df[r1,i]) #append 
 #pivotted rows to columns of the tourn_all df

}

#Select only needed columns
col_index <- c(1:10, 12:13)
tourn_all_data <- tourn_all_sub[,col_index]


#set row 1 as column names and delete the row
colnames(tourn_all_data) <- tourn_all_data[1,]
tourn_all_data <- tourn_all_data[-1,]


Row_no <- nrow(tourn_all_data)

stage_str <-  vector(mode="character", length=Row_no)

for (i in 1:Row_no)
{
  stage_str[i] <- paste (tourn_all_data[i, 4], tourn_all_data[i, 5],   tourn_all_data[i, 6],tourn_all_data[i, 7],tourn_all_data[i, 8],tourn_all_data[i, 9],tourn_all_data[i, 10],sep = " ", collapse = NULL)

}

tourn_all_data[, 13] <-
str_count(stage_str, '[W]')


tourn_all_data[, 14] <-
str_count(stage_str, '[D]')

tourn_all_data[, 15] <-
str_count(stage_str, '[L]')



colnames(tourn_all_data) <- c("Pair Number", "Player Name", "Total", "Round 1", "Round 2", "Round 3", "Round 4", "Round 5", "Round 6", "Round 7", "Player State", "Player Pre-Ranking","Number of Wins", "Number of Draws","Number of Loses")

#View(tourn_all_data)
```



#Pre vs Post Tournament Rankings

##Collect Post Tournament Rankings
```{r post-rankings, eval = TRUE}
extra_data <- integer(0)
j <- 1
k <- 1

while( j < length(raw_data)) {

  postranking_info <- str_replace_all(raw_data[j+12], " ","")
  post_rank <- as.numeric(str_replace(str_extract(postranking_info, "[>][[:digit:]]{3,4}"), ">", ""))
  extra_data[k] <- post_rank
  k <- k + 1
  j <- j +23
} 

newtournamentData <- tournament_data
newtournamentData[ , "Post Ratings"] <- extra_data

```



##Scatterplot show pre vs post torunament ratings

```{r scatter-plot, eval = TRUE}
plot(newtournamentData$`Player's Pre Rating`, newtournamentData$`Post Ratings`)
```


The scatterplot shows that for the most part there is positive correlation between the pre and post tournament rankings - as expected.



## Add number of wins, losses, and draws to dataframe

```{r}
newtournamentData[ , c('Number of Wins','Number of Draws', 'Number of Loses')] <- tourn_all_data[,c('Number of Wins','Number of Draws', 'Number of Loses')]

```


#Ranking players with most wins, draws and loss 

```{r}


head(newtournamentData[order(-newtournamentData$`Number of Wins`),c(1,7)])


head(newtournamentData[order(-newtournamentData$`Number of Draws`),c(1,8)])


head(newtournamentData[order(-newtournamentData$`Number of Loses`),c(1,9)])
```
#Histogram of number wins 


```{r}
hist(newtournamentData$`Number of Wins`)
```
#Histogram of number Draws
```{r}
hist(newtournamentData$`Number of Draws`)
```
#Histogram of number Loses
```{r}
hist(newtournamentData$`Number of Loses`)
```